Reanalysis Step 2: Multiple Imputation (Clean Version)

### 2025-08-29 reanalysis-2-multiple_imputation.qmd
more predictor

# Variables that should always be included as predictors (key variables)
vars_always_predict <- c(
  "ever_lapse_binary",      # Exposure
  "baseline_VA_logMAR",     # Key outcome
  "outcome_VA_logMAR",      # Key outcome
  "age",                    # Demographics
  "gender_cat",
  "race_ethnic_cat",        # Demographics
  "insurance_cat",          # Demographics/SES
  "CCI", "DCSI",           # Comorbidities
  "person_dr",              # DR severity - key clinical variable
  "anti_VEGF",              # Key treatment
  "focal_laser_flag",       # Treatment
  "PRP_flag"                # Treatment
)

@@

# Comorbidities should predict DR severity and vice versa
if ("person_dr" %in% rownames(pred_matrix)) {
  if ("CCI" %in% colnames(pred_matrix)) pred_matrix["person_dr", "CCI"] <- 1
  if ("DCSI" %in% colnames(pred_matrix)) pred_matrix["person_dr", "DCSI"] <- 1
  if ("glaucoma_bef_hitplus_cat" %in% colnames(pred_matrix)) pred_matrix["person_dr", "glaucoma_bef_hitplus_cat"] <- 1
}

@@

# Fellow-eye predictors with weight=2 for stronger influence
fellow_vars <- c("fellow_baseline_VA", "fellow_intermediate_VA", "fellow_outcome_VA", "fellow_dr")
for (v in fellow_vars) {
  if (v %in% rownames(pred_matrix)) {
    pred_matrix[v, ] <- 0
  }
}
if ("fellow_baseline_VA" %in% colnames(pred_matrix)) {
  pred_matrix["baseline_VA_logMAR", "fellow_baseline_VA"] <- 1
}
if ("fellow_intermediate_VA" %in% colnames(pred_matrix)) {
  pred_matrix["intermediate_VA_logMAR", "fellow_intermediate_VA"] <- 1
}
if ("fellow_outcome_VA" %in% colnames(pred_matrix)) {
  pred_matrix["outcome_VA_logMAR", "fellow_outcome_VA"] <- 1
}
if ("fellow_dr" %in% colnames(pred_matrix) && "person_dr" %in% rownames(pred_matrix)) {
  pred_matrix["person_dr", "fellow_dr"] <- 1
}

>>>>

# Variables that should always be included as predictors (key variables)
vars_always_predict <- c(
  "ever_lapse_binary",      # Exposure
  "baseline_VA_logMAR",     # Key outcome
  "outcome_VA_logMAR",      # Key outcome
  "age",                    # Demographics
  "gender_cat",
  "race_ethnic_cat",        # Demographics
  "insurance_cat",          # Demographics/SES
  "CCI", "DCSI",           # Comorbidities
  "person_dr",              # DR severity - key clinical variable
  "anti_VEGF",              # Key treatment
  "focal_laser_flag",       # Treatment
  "PRP_flag",               # Treatment
  "glaucoma_bef_hitplus_cat",  # Ocular comorbidity
  "otherretina_bef_hitplus_cat", # Ocular comorbidity
  "catsurg_before_hitplus_cat",  # Cataract surgery
  "fellow_baseline_VA", "fellow_intermediate_VA", "fellow_outcome_VA", "fellow_dr" # Fellow-eye predictors
)

@@

# Comorbidities should predict DR severity and vice versa
if ("person_dr" %in% rownames(pred_matrix)) {
  if ("CCI" %in% colnames(pred_matrix)) pred_matrix["person_dr", "CCI"] <- 1
  if ("DCSI" %in% colnames(pred_matrix)) pred_matrix["person_dr", "DCSI"] <- 1
  if ("glaucoma_bef_hitplus_cat" %in% colnames(pred_matrix)) pred_matrix["person_dr", "glaucoma_bef_hitplus_cat"] <- 1
  if ("otherretina_bef_hitplus_cat" %in% colnames(pred_matrix)) pred_matrix["person_dr", "otherretina_bef_hitplus_cat"] <- 1
  if ("catsurg_before_hitplus_cat" %in% colnames(pred_matrix)) pred_matrix["person_dr", "catsurg_before_hitplus_cat"] <- 1
  if ("fellow_dr" %in% colnames(pred_matrix)) pred_matrix["person_dr", "fellow_dr"] <- 1
}

@@

# Fellow-eye predictors: allow them to predict VA and DR
fellow_vars <- c("fellow_baseline_VA", "fellow_intermediate_VA", "fellow_outcome_VA", "fellow_dr")
for (v in fellow_vars) {
  if (v %in% rownames(pred_matrix)) {
    pred_matrix[v, ] <- 0
  }
}
if ("fellow_baseline_VA" %in% colnames(pred_matrix)) pred_matrix["baseline_VA_logMAR", "fellow_baseline_VA"] <- 1
if ("fellow_intermediate_VA" %in% colnames(pred_matrix)) pred_matrix["intermediate_VA_logMAR", "fellow_intermediate_VA"] <- 1
if ("fellow_outcome_VA" %in% colnames(pred_matrix)) pred_matrix["outcome_VA_logMAR", "fellow_outcome_VA"] <- 1
if ("fellow_dr" %in% colnames(pred_matrix) && "person_dr" %in% rownames(pred_matrix)) pred_matrix["person_dr", "fellow_dr"] <- 1

# Ocular comorbidities and cataract surgery predict VA timepoints
for (va in c("baseline_VA_logMAR", "intermediate_VA_logMAR", "outcome_VA_logMAR")) {
  for (ov in c("glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat", "catsurg_before_hitplus_cat")) {
    if (va %in% rownames(pred_matrix) && ov %in% colnames(pred_matrix)) pred_matrix[va, ov] <- 1
  }
}

# Exposure predicts VA timepoints (lapse history)
for (va in c("baseline_VA_logMAR", "intermediate_VA_logMAR", "outcome_VA_logMAR")) {
  if (va %in% rownames(pred_matrix) && "ever_lapse_binary" %in% colnames(pred_matrix)) pred_matrix[va, "ever_lapse_binary"] <- 1
}
